# (PART) HR Analytics practical examples{-}

# Gender Pay Gap {#pay-gap}

```{r pay-gap-analysis, include=FALSE}
chap <- 13
lc <- 0
rq <- 0
# **`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**
# **`r paste0("(RQ", chap, ".", (rq <- rq + 1), ")")`**

knitr::opts_chunk$set(
  tidy = FALSE, 
  out.width = '\\textwidth', 
  fig.height = 4,
  warning = FALSE
  )

options(scipen = 99, digits = 3)

# Set random number generator see value for replicable pseudorandomness. Why 76?
# https://www.youtube.com/watch?v=xjJ7FheCkCU
set.seed(76)
```


According to a recent Glassdoor survey, more than two-thirds (67 percent)
of U.S. employees say they would not apply for jobs at employers where they
believe a gender pay gap exists.1 Today, the gender pay gap is more than
a social or legal issue. It’s an issue that can affect the ability of employers to
attract and retain talent.

How should HR practitioners react to concerns about the gender pay gap?
One increasingly popular way is to perform an internal gender pay audit to
understand whether a gap exists at your company. This involves examining
your own payroll data for evidence of a gender pay gap, and making
recommendations to senior management about ways to lower gender
barriers in recruitment, hiring, pay and promotion before they arise as
broader organizational concerns.

The following example appeared on the Glassdor website in March 2017, with the title "How to Audit Your Gender Pay Gap: An Employers Guide Using R" and was written by Andrew Chamberlain, Ph.D.


```{r packages}
# Load R libraries.
library(tidyverse)
library(tidymodels)
library(devtools)
library(HRAnalytics) # Helper package

#Turn off scientific notation. 
options(scipen = 999)
``` 

```{r load data}
# Load data. 
gd_data = read_csv("https://glassdoor.box.com/shared/static/beukjzgrsu35fqe59f7502hruribd5tt.csv")
# N = 1000 total observations.
```


## Data Cleaning and Prep.

Before we can analyse whether we have gender pay gap, we need to prepare our data so it is cleaned up and has useful columns we can analyse.

We will perform the following steps in our process:

1. Make a number of sensible age brackets e.g. <25, 25-34, 35-44, 45-54, >54
2. Create total compensation variable (base pay + bonus)
3. Take natural logarithm of compensation variables (for percentage pay gap interpretation in regressions).
4. Convert values like department into values that can act as flags in our model and making the most commonly occuring value the default which everything else gets compared to in our model.

```{r Summary Statistics}
gd_data %>%
  # Age brackets
  mutate(age_bin = cut(age,
                  breaks = c(0, 25, 35, 45, 55, Inf),
                  right = FALSE)) %>%
  # Total compensation
  mutate(total_pay = basePay + bonus) %>%
  # Log of compensation
  mutate(log_base = log(basePay, base = exp(1)),
         log_total = log(total_pay, base = exp(1)),
         # Adds 1 to allow for log of 0 bonus values.
         log_bonus = log(bonus + 1, base = exp(1))) %>%
  # Make flags
  mutate_if(is_character, fct_infreq) %>%
  mutate(age_bin = fct_infreq(age_bin)) ->
  gd_data_clean
```


## Summary Statistics by gender
We can look at summary statistics by gender.

```{r gd_summary_gender_base}
# Base pay summary stats.
gd_data_clean %>%
  # Exclude stuff with missing info
  filter(!is.na(basePay)) %>%
  # Analyse by gender
  group_by(gender) %>%
  # Retrieve summary statistics
  summarise(
    mean_base = mean(basePay),
    median_base = median(basePay),
    count = n()
  ) ->
  gd_summary_gender_base

gd_summary_gender_base
```

```{r gd_summary_gender_total}
# Total pay summary stats.
gd_data_clean %>%
  # Use a helper function to save typing
  HRAnalytics::gap_summary(gender, total_pay) ->
  gd_summary_gender_total

gd_summary_gender_total
```

```{r gd_summary_gender_bonus}
# Bonus summary stats. 
gd_data_clean %>%
  # Use a helper function to save typing
  HRAnalytics::gap_summary(gender, bonus) ->
  gd_summary_gender_bonus

gd_summary_gender_bonus
```

```{r gd_summary_gender_perf}
# Performance evaluations summary stats. 
gd_data_clean %>%
  # Use a helper function to save typing
  HRAnalytics::gap_summary(gender, perfEval)  ->
  gd_summary_gender_perf

gd_summary_gender_perf
```

## Avoiding Simpson's Paradox
Simspson's Paradox is where some high level aggregations lead to a different conclusion than would have been arrived at if a lower level of detail had been looked at. We should check out values split by department and job title to see if they showcase wage differences.

```{r gd_summary_dept_gender_total}
# Performance evaluations summary stats. 
gd_data_clean %>%
  filter(!is.na(total_pay)) %>%
  # Create a summary by department and gender
  group_by(dept, gender) %>%
  summarise(
    mean_perf = mean(total_pay),
    median_perf = median(total_pay),
    count = n()
  ) %>%
  # "Unpivot" the data
  gather(measure, value, mean_perf:count) %>%
  # Combine the gender with measure
  unite(combo, measure, gender) %>%
  # "Pivot" the data to see all the measures split by gender
  spread(combo, value) ->
  gd_summary_dept_gender_total

gd_summary_dept_gender_total
```

```{r gd_summary_job_gender_total}
# Performance evaluations summary stats. 
gd_data_clean %>%
  filter(!is.na(total_pay)) %>%
  # Create a summary by job and gender
  group_by(jobTitle, gender) %>%
  summarise(
    mean_perf = mean(total_pay),
    median_perf = median(total_pay),
    count = n()
  ) %>%
  # "Unpivot" the data
  gather(measure, value, mean_perf:count) %>%
  # Combine the gender with measure
  unite(combo, measure, gender) %>%
  # "Pivot" the data to see all the measures split by gender
  spread(combo, value) ->
  gd_summary_job_gender_total

gd_summary_job_gender_total
```

## Model Estimation: OLS with controls. 
Coefficient on "male" has the interpretation of approximate male pay advantage ("gender pay gap").


### Logarithm of Base Pay


```{r Linear models}
# No controls. ("unadjusted" pay gap.)
lm_gender <- lm(log_base ~ gender, data = gd_data_clean)

# Adding "human capital" controls (performance evals, age and education).
lm_humancapital <- lm(log_base ~ gender + perfEval + age_bin + edu, data = gd_data_clean)

# Adding all controls. ("adjusted" pay gap.)
lm_allcontrols <- lm(log_base ~ gender + perfEval + age_bin + edu + dept + seniority + jobTitle, data = gd_data_clean)
```

```{r lm_gender}
lm_gender %>%
  summary()
```

```{r lm_gender viz}
lm_gender %>%
  # Get the predicted values
  augment() %>%
  # Rename columns to more friendly names
  rename(actual = log_base, predicted = .fitted) %>%
  # Build a chart
  ggplot() +
  # Add columns to it
  aes(x=actual, y=predicted) +
  # Choose chart type
  geom_point() +
  # Add a diagonal line representing perfect predictions
  geom_abline(colour="blue", slope=1, intercept=0)+
  # Split by gender
  facet_wrap(~gender) +
  # Add some labels
  labs(title="Actual vs predicted", 
       subtitle="Values predicted using a linear model containing gender")
```

```{r lm_humancapital}
lm_humancapital %>% 
  summary()
```


```{r lm_humancapital viz}
lm_humancapital %>% 
  # Get the predicted values
  augment() %>% 
  # Rename columns to more friendly names
  rename(actual = log_base, predicted = .fitted) %>% 
  HRAnalytics::pred_vs_actuals() +
  # Add some labels
  labs(title="Actual vs predicted", 
       subtitle="Values predicted using a linear model containing human capital measures")
```

```{r lm_allcontrols}
lm_allcontrols %>% 
  summary()
```

```{r lm_allcontrols viz}
lm_allcontrols %>% 
  # Get the predicted values
  augment() %>% 
  # Rename columns to more friendly names
  rename(actual = log_base, predicted = .fitted) %>% 
  HRAnalytics::pred_vs_actuals() +
  # Add some labels
  labs(title="Actual vs predicted", 
       subtitle="Values predicted using a linear model all controls")
```

```{r coefficients}
#  Gather up all the models
list(lm_gender, lm_humancapital, lm_allcontrols) %>% 
  # Extract coefficients for all models at once and combine into a single table
  map_df(tidy, .id = "model") %>% 
  # Let's look at the impact of gender
  filter(term=="genderFemale") %>%
  # P values less than 0.05 are usually taken to mean an estimate is reliable
  select(model, log_gap=estimate, p.value)
```

Note that too little data can make results unreliable for a complex model because there are so few records to use for each combination of values.

### Results by Department
(Interaction of gender x dept)
To test for differences by department, examine significance of each "gender x dept" coefficient.
For the gender pay gap by department, add the "gender" + "gender x dept" coefficients from this model. 

```{r lm_allcontrols_dept}
# All controls with department interaction terms. 
lm_allcontrols_dept <- lm(log_base ~ gender*dept + perfEval + age_bin + edu + seniority + jobTitle, data = gd_data_clean)

tidy(lm_allcontrols_dept)

lm_allcontrols_dept %>% 
  # Get the predicted values
  augment() %>% 
  # Rename columns to more friendly names
  rename(actual = log_base, predicted = .fitted) %>% 
  HRAnalytics::pred_vs_actuals() +
  # Add some labels
  labs(title="Actual vs predicted", 
       subtitle="Values predicted using a linear model all controls & department interaction")
```

### Results by Job Title 
(Interaction of gender x job )
To test for differences by department, examine significance of each "gender x job title" coefficient.
For the gender pay gap by job, add the "gender" + "gender x job title" coefficients from this model. 


```{r lm_allcontrols_job}
# All controls with department interaction terms. 
lm_allcontrols_job <- lm(log_base ~ gender*jobTitle + perfEval + age_bin + edu + seniority + dept, data = gd_data_clean)

tidy(lm_allcontrols_job)

lm_allcontrols_job %>% 
  # Get the predicted values
  augment() %>% 
  # Rename columns to more friendly names
  rename(actual = log_base, predicted = .fitted) %>% 
  # Build a chart
  HRAnalytics::pred_vs_actuals() +
  # Add some labels
  labs(title="Actual vs predicted", 
       subtitle="Values predicted using a linear model all controls & job interaction")
```

For additional analysis via Oaxaca-Blinder decomposition, please see documentation for the [`oaxaca` package](https://cran.r-project.org/package=oaxaca) in R.
